### 2019-02-12 coursework/experiment-design/design.tex

\documentclass{article}
\usepackage[a4paper]{geometry}
\title{COMP80131 Assignment 1\\Dynamic Strategy Priority for Vampire}
\author{Michael Rawson}
\begin{document}
\maketitle

\section{Task}
Modern automated theorem provers (e.g. Vampire~\cite{Vampire}) use a fixed \emph{portfolio} of strategies~\cite{portfolio} to direct search for a proof, of which only a few might solve a hard problem quickly.
Portfolio provers execute their strategies in sequence, with a pre-defined order, and a time allocation for each strategy.
However, it is not known whether a strategy will succeed, and is is often the case that a failing strategy is left to churn for considerable time without interruption~\cite{predict-success}.
We investigate the feasibility of training a machine-learned classifier to predict the success or failure of a strategy from data drawn from an initial period of its execution.

\section{Rationale}
If it proves possible to classify running strategies, it will then allow the construction of a \emph{portfolio scheduler} which defers execution of strategies the classifier deems less likely to succeed until after more-promising strategies have run.
In this way, Vampire will find proofs faster on average, theoretically without losing any proofs it found previously.
The experiment will answer ``can one of a (fixed) list of machine-learning methods produce a statistically-significant classifier for this problem'' --- note that no fixed accuracy is required, as any working classifier will improve the current state of Vampire.
We hypothesise that at least one method will beat chance.
The expectation is that most methods will succeed in this, but that no classifier will achieve a high level of accuracy, as intuitively this is a ``hard problem'', similar to that of the undecidable \emph{Entscheidungsproblem}~\cite{turing}.

\section{Methodology}
The experiment will measure the performance of a fixed set of supervised classification methods.
Classifiers will be trained on a balanced set of Vampire strategy execution traces, tagged with the strategy's eventual success or failure.
Performance may then be evaluated by a variety of statistical methods, but all involve using the trained classifier with an ``unseen'' or ``test'' set (to avoid evaluating the classifier's tendency to overfit~\cite{overfitting}) and observing the number of correct vs. incorrect classifications.

\section{Data}
The training (and test) data will be obtained by running a modified version of Vampire on a set of problems~\cite{TPTP} designed for the CASC~\cite{CASC} benchmark.
This version will log execution data and the success or failure of strategies.
Since there are a large number of strategies and many problems, sufficient data may be obtained easily.

\section{Repetition}
Vampire is now open-source, and results from this experiment will be published.
Therefore, repetition by other researchers is possible and encouraged.
Data will be available, but may also be recreated without significant effort.

\section{Analysis}
Classifier performance will be evaluated by means of \(k\)-fold cross-validation~\cite{cross-validation}.
This will allow for both reliable estimates of classifier accuracy from all available data, and also calculation of statistical significance tests such as a \(p\)-value from cross-validation results.
If a classifier exceeds the null hypothesis with high statistical significance, the hypothesis may be accepted.

\bibliographystyle{plain}
\bibliography{design.bib}
\end{document}